---
title: "Homework 8"
author: "Kavya Malgi"
date: '`r Sys.Date()`'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      eval =TRUE, 
                      warning = FALSE,
                      message = FALSE,
                      include = TRUE,
                      fig.align = "center",
                      fig.show='hold',
                      R.options = list(max.print=50))


library(tibble)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(rvest)
library(sas7bdat)
library(knitr)
library(MatchIt)
library(boot)
```

Github Link: <https://github.com/kavyamalgi/Homework8>

## Problem 1: Regression Warm Up

**A) What creatinine clearance rate should we expect for a 55 year-old? Explain briefly (one or two sentences + equations) how you determined this.**

```{r}

creat <- read_csv("creatinine.csv")

creat_model <- lm(creatclear ~ age, data = creat)
#summary(creat_model)
```

After running the regression model equation, I calculated the intercept as 147.81292 and the slope as -0.61982. Using this we can say our equation is:

**creatclear = 147.81292 - 0.61982(age)**

For a **55 year old man**:

creatclear(55 years) = **113.72**

So based on the fitted linear model, we'd expect a 55 year old to have a creatinine clearance rate of 113.72 mL/min per year

------------------------------------------------------------------------

**B) How does creatinine clearance rate change with age? (This should be a single number whose units are ml/minute per year.) Explain briefly (one or two sentences) how you determined this.**

**The creatinine clearance rate decreases by 0.61982 mL/min per year for each additional year of age.**

I determined this by looking at my regression output from the summary(creat_model) and looked at the coefficient for age.

------------------------------------------------------------------------

**C) Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? Explain briefly (a few sentences + equations) how you determined this.**

Using the regression equation:

[**40 year old:**]{.underline}

147.81292 - 0.61982(40) = **123.02012**

[**60 year old:**]{.underline}

147.81292 - 0.61982(60) = **110.62372**

------------------------------------------------------------------------

\newpage

[**Compare actual - expected:**]{.underline}

[**40 year old:**]{.underline}

135 - 123.02012 = **11.97988**

[**60 year old:**]{.underline}

112 - 110.62372 = **1.37628**

After calculating the expected using the regression equation and comparing the actual (given) - expected (calculated), the 40 year old's creatinine clearance rate is farther above the expected value by about 12 mL/min per year, compared to the 60 year old who is only about 1.4 above the expected value.

**Therefore we can conclude that the 40 year old has the healthier creatinine clearance rate for their age.**

------------------------------------------------------------------------

## Problem 2: Modeling Disease Growth

**1) An estimated growth rate and doubling time for Italy, with 95% bootstrapped confidence intervals for each.**

```{r}
covid <- read_csv("covid.csv")

italy <- covid |>
  filter(country == "Italy", deaths >0) |>
  mutate(log_deaths = log(deaths))

mod_i <- lm(log_deaths ~ days_since_first_death, data = italy)

italy_growth <- coef(mod_i)[2]
italy_double <- log(2) / italy_growth


bootstrap_italy <- function(data, indices) {
  d <- data[indices, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = d)
  growth <- coef(model)[2]
  doubling <- log(2) / growth
  return(c(growth, doubling))
}

set.seed(123)
boot_i <- boot(data = italy, statistic = bootstrap_italy, R = 1000)

italy_ci <- boot.ci(boot_i, type = "perc", index = 1)
italy_ci <- boot.ci(boot_i, type = "perc", index = 2)
italy_ci_growth <- boot.ci(boot_i, type = "perc", index = 1)
#print(italy_ci_growth)

italy_ci_doubling <- boot.ci(boot_i, type = "perc", index = 2)
#print(italy_ci_doubling)




```

**Estimated Growth Rate for Italy:** 0.183/day

**Growth Rate CI for Italy:** (0.159, 0.208)

**Estimated Doubling Time for Italy:** 3.8 days for daily deaths to double

**Doubling Time CI for Italy:** ( 3.3, 4.4)

------------------------------------------------------------------------

**2) An estimated growth rate and doubling time for Spain, with 95% bootstrapped confidence intervals for each.**

```{r}
spain <- covid |>
  filter(country == "Spain", deaths >0) |>
  mutate(log_deaths = log(deaths))

mod_s <- lm(log_deaths ~ days_since_first_death, data = spain)

growth_s <- coef(mod_s)[2]
double_s <- log(2) / growth_s

bootstrap_spain <- function(data, indices) {
  d <- data[indices, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = d)
  growth <- coef(model)[2]
  doubling <- log(2) / growth
  return(c(growth, doubling))
}

set.seed(123)
spain_boot <- boot(data = spain, statistic = bootstrap_spain, R = 1000)

spain_ci_grow <- boot.ci(spain_boot, type = "perc", index = 1)
spain_ci_double <- boot.ci(spain_boot, type = "perc", index = 2)


#print(spain_ci_grow)


#print(spain_ci_double)

```

**Estimated Growth Rate for Spain** 0.276/day

**Growth Rate CI for Spain:** (0.233, 0.319)

**Estimated Doubling Time for Spain:** 2.5 days for daily deaths to double

**Doubling Time CI for Spain:** ( 2.2, 3.0)

------------------------------------------------------------------------

**3) A line graph showing reported daily deaths over time (using days_since_first_death, rather than calendar date, as the relevant time variable) in each country. Your line graph should have two lines, one for each country, distinguished by their color.**

```{r}


ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(size = 1) +
  labs(
    title = "Reported Daily COVID-19 Deaths by Day Since 1st Death",
    x = "Days Since First Death",
    y = "Daily Reported Deaths",
    color = "Country"
  )


```

------------------------------------------------------------------------

## Problem 3: Price Elasticity of Demand

**In light of the data, what is the estimated price elasticity of demand for milk? Give a 95% bootstrapped confidence interval for this quantity. Briefly describe what you did to estimate the elasticity – no more than a few sentences, together with your estimate + interval.\
**

```{r}
milk <- read_csv("milk.csv")
milk <- milk |>
  mutate(log_price = log(price),
         log_sales = log(sales))

elasticity_model <- lm(log_sales ~ log_price, data = milk)
#summary(elasticity_model)

price_elasticity <- coef(elasticity_model)[2]

boot_milk <- function(data, indices) {
  d<- data[indices, ]
  model <- lm(log_sales ~ log_price, data = d)
  return(coef(model)[2])
}

set.seed(123)
milk_boot <- boot(data = milk, statistic = boot_milk, R = 1000)

#elasticity_CI <- boot.ci(milk_boot, type = "perc")



```

**log(sales) = intercept + slope(log(price))**

**The Estimated Elasticity: -1.619**

I estimated the price elasticity of demand for milk by using a log-log linear regression of quantity purchased on price. This means that a 1% increase in price leads to a 1.62% decrease in quantity demand, meaning elastic demand. I then found the 95% bootstrapped confidence interval as (-1.771, -1.439). This means that we are 95% confident that the true price elasticity of demand for milk lies between the interval.
